---
title: "Alexia"
author: "Marialexia Alfano"
date: "October 30, 2017"
output: html_document
---

ggcorrplot & ggbeeswarm 

ggbeeswarm [vignette](https://cran.r-project.org/web/packages/ggbeeswarm/vignettes/usageExamples.pdf)

ggcorrplot [vignette](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2) 

#ggbeeswarm

```{r}
library(ggbeeswarm)
set.seed(12345)
n<-100
dat<-rnorm(n*2)
labs<-rep(c('a','b'),n)
ggplot(mapping=aes(labs, dat)) + geom_quasirandom()
```

```{r}
ggplot(mapping=aes(labs, dat)) + geom_quasirandom(aes(color=labs))
```


```{r}
labs2<-factor(labs,levels=c('b','a'))
ggplot(mapping=aes(labs2, dat)) + geom_quasirandom(aes(color=labs))
```


```{r}
ggplot(mapping=aes(dat,labs)) + geom_quasirandom(aes(color=labs),groupOnX=FALSE)
```

```{r}
labs2<-factor(rep(1:2,each=n))
ggplot(mapping=aes(labs,dat,color=labs2)) + geom_quasirandom(dodge.width=.8)
```

```{r}
labs2<-factor(rep(1:2,each=n))
ggplot(mapping=aes(dat,labs,color=labs2)) + geom_quasirandom(dodge.width=.8,groupOnX=FALSE)
```

```{r}
ggplot(mapping=aes(labs,dat,color=labs2))+
geom_beeswarm(dodge.width=.8,cex=2)
```

```{r}
ggplot(mapping=aes(dat,labs,color=labs2))+
geom_beeswarm(dodge.width=.8,cex=2,groupOnX=FALSE)

```

```{r}
 df<-data.frame(labs,dat,labs2)
ggplot(df,aes(labs,dat,color=labs2))+
geom_quasirandom()+
facet_grid(.~labs2)

```

```{r}
ggplot(df,aes(labs,dat,color=labs2))+
geom_beeswarm(cex=3)+
facet_grid(.~labs2)

```

```{r}
library(gridExtra)
library(ggplot2)
library(ggthemes)
dat <- list('Normal'=rnorm(50),'Dense normal'= rnorm(500),'Bimodal'=c(rnorm(100), rnorm(100,5)),'Trimodal'=c(rnorm(100), rnorm(100,5),rnorm(100,-3)))

labs<-rep(names(dat),sapply(dat,length))
labs<-factor(labs,levels=unique(labs)) 
dat<-unlist(dat)

   p1<-ggplot(mapping=aes(labs, dat))+
     geom_beeswarm(alpha=.2,size=.75)+
     ggtitle('geom_beeswarm') + labs(x='')+
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
   
grid.arrange(p1)
```

 






#ggcorrplot

```{r}
library(ggcorrplot)
data(mtcars)
corr <- round(cor(mtcars), 1)
head(corr[, 1:6])
```

```{r}
p.mat <- cor_pmat(mtcars)
head(p.mat[, 1:4])
```

```{r}
ggcorrplot(corr)
```

```{r}
ggcorrplot(corr, method = "circle")
```

####reordering the correlation matrix using heirarchical clustering
```{r}
ggcorrplot(corr, hc.order = TRUE, outline.col = "white")
```


####show lower triangle
```{r}
ggcorrplot(corr, hc.order = TRUE, type = "lower",
     outline.col = "white")
```

####show upper triangle
```{r}
ggcorrplot(corr, hc.order = TRUE, type = "upper",
     outline.col = "white")
```

####change colors 
```{r}
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   colors = c("#6D9EC1", "white", "#E46726"))
```

####add correlation coefficients in squares
####argument is lab = TRUE
```{r}
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)
```

####Add correlation significance level
####Argument p.mat
####Barring the no significant coefficient
```{r}
ggcorrplot(corr, hc.order = TRUE,
    type = "lower", p.mat = p.mat)
```

####Leave blank on no significant coefficient
```{r}
ggcorrplot(corr, p.mat = p.mat, hc.order = TRUE,
    type = "lower", insig = "blank")
```

