---
title: "Colleen"
author: "Colleen Brents"
date: "October 30, 2017"
output:
  word_document: default
  html_document: default
---
clean code is from [here](http://www.github.com/KatieKey/input_output_shiny_group/CSV_Files)
focus on getting stuff to run; bee swarm, heirarchical, correlation ; 3 functions 
alluvial - compare against the control population (are the parameters for the drug concentrations outside the 95% CI of the control) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the Rmd for ggalluvial and a bonus 'exploratory' package for the group 2 project 
[link for ggplot extnesions](http://www.ggplot2-exts.org/) 
[actual link](http://www.ggplot2-exts.org/gallery/)
here is the link for ggalluvial [here](https://github.com/corybrunson/ggalluvial)
here's a blog using other packages [here](https://www.r-bloggers.com/alluvial-diagrams/)

```{r}
devtools::install_github("corybrunson/ggalluvial", build_vignettes = TRUE)
library(ggplot2)
library(ggalluvial)
```
now that we've downloaded the packages and libraries, check out a vignette:

```{r}
vignette(topic = "ggalluvial")
```
sample data is Titanic 


Other packages people in group2 are looking at : 
+ heirarchical clustering
+ ggraph
+ geomnet
+ ggalluvial
+ ggmosaic
+ ggcorrplot
+ 'geom_tile' --> heatmaps 
  
#if publish as html, opens the vignette
#if publish as word, opens my comments 
  + a bonus ggplot 


Run the test data from th vignette 
```{r, run the test data}
library(ggplot2)
library(ggalluvial)
```

```{r, UC berkley data}
head(as.data.frame(UCBAdmissions), n = 12)
is_alluvial(as.data.frame(UCBAdmissions), logical = FALSE)
ggplot(as.data.frame(UCBAdmissions),
       aes(weight = Freq, axis1 = Gender, axis2 = Dept)) +
  geom_alluvium(aes(fill = Admit), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", label.strata = TRUE) +
  scale_x_continuous(breaks = 1:2, labels = c("Gender", "Dept")) +
  ggtitle("UC Berkeley admissions and rejections, by sex and department")
```

```{r, Titanic data}
ggplot(as.data.frame(Titanic),
       aes(weight = Freq,
           axis1 = Survived, axis2 = Sex, axis3 = Class)) +
  geom_alluvium(aes(fill = Class),
                width = 0, knot.pos = 0, reverse = FALSE) +
  guides(fill = FALSE) +
  geom_stratum(width = 1/8, reverse = FALSE) +
  geom_text(stat = "stratum", label.strata = TRUE, reverse = FALSE) +
  scale_x_continuous(breaks = 1:3, labels = c("Survived", "Sex", "Class")) +
  coord_flip() +
  ggtitle("Titanic survival by class and sex")
```
could be used for showing biological or chemical markers; have to make categorical  

```{r, one row per lode}
UCB_lodes <- to_lodes(as.data.frame(UCBAdmissions), axes = 1:3)
head(UCB_lodes, n = 12)
is_alluvial(UCB_lodes, logical = FALSE)
```
```{r, weighted things }
data(Refugees, package = "alluvial")
country_regions <- c(
  Afghanistan = "Middle East",
  Burundi = "Central Africa",
  `Congo DRC` = "Central Africa",
  Iraq = "Middle East",
  Myanmar = "Southeast Asia",
  Palestine = "Middle East",
  Somalia = "Horn of Africa",
  Sudan = "Central Africa",
  Syria = "Middle East",
  Vietnam = "Southeast Asia"
)
#instead of country regions make chemical or biological markers 
#drugdose on x axis
#this is for continuous data 
Refugees$region <- country_regions[Refugees$country]
ggplot(data = Refugees,
       aes(x = year, weight = refugees, alluvium = country)) +
  geom_alluvium(aes(fill = country, colour = country),
                alpha = .75, decreasing = FALSE) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~ region, scales = "fixed") +
  ggtitle("refugee volume by country of origin and geographic region")
```

```{r, }
data(majors)
majors$curriculum <- as.factor(majors$curriculum)
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum, label = curriculum)) +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft") +
  geom_stratum() +
  theme(legend.position = "bottom") +
  ggtitle("student curricula across several semesters")
```

```{r}
data(majors)
majors$curriculum <- as.factor(majors$curriculum)
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum, label = curriculum)) +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft") +
  geom_stratum() +
  theme(legend.position = "bottom") +
  ggtitle("student curricula across several semesters")
```

