---
title: "Colleen"
author: "Colleen Brents"
date: "October 30, 2017"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the Rmd for ggalluvial and a bonus 'exploratory' package for the group 2 project 
[link for ggplot extnesions](http://www.ggplot2-exts.org/) 
[actual link](http://www.ggplot2-exts.org/gallery/)
here is the link for ggalluvial [here](https://github.com/corybrunson/ggalluvial)
here's a blog using other packages [here](https://www.r-bloggers.com/alluvial-diagrams/)

```{r}
devtools::install_github("corybrunson/ggalluvial", build_vignettes = TRUE)
library(ggplot2)
library(ggalluvial)
```
now that we've downloaded the packages and libraries, check out a vignette:

```{r}
vignette(topic = "ggalluvial")
```
sample data is Titanic 


Other packages people in group2 are looking at 
  + heirarchical clustering
  + ggraph
  + geomnet
  + ggalluvial
  + ggmosaic
  + ggcorrplot
  + 'geom_tile' --> heatmaps 
  
#if publish as html, opens the vignette
#if publish as word, opens my comments 
  + a bonus ggplot 
