---
title: "Mackenzie_Extension"
author: "Mackenzie Fry"
date: "October 30, 2017"
output: word_document
---



```{r global_options}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
warning = FALSE)
```



#GGRAPH package

[link to ggraph package website](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html) 
 

Packages needed for this method: **ggraph, igraph**

##Layouts   



Here is the sample code:  
```{r}
library(ggraph)
library(igraph)
set_graph_style(plot_margin = margin(1,1,1,1))
hierarchy <- as.dendrogram(hclust(dist(iris[, 1:4])))

# Classify nodes based on agreement between children
hierarchy <- tree_apply(hierarchy, function(node, children, ...) {
    if (is.leaf(node)) {
        attr(node, 'Class') <- as.character(iris[as.integer(attr(node, 'label')),5])
    } else {
        classes <- unique(sapply(children, attr, which = 'Class'))
        if (length(classes) == 1 && !anyNA(classes)) {
            attr(node, 'Class') <- classes
        } else {
            attr(node, 'Class') <- NA
        }
    }
    attr(node, 'nodePar') <- list(class = attr(node, 'Class'))
    node
}, direction = 'up')

hairball <- graph_from_data_frame(highschool)

# Classify nodes based on popularity gain
pop1957 <- degree(delete_edges(hairball, which(E(hairball)$year == 1957)), 
                  mode = 'in')
pop1958 <- degree(delete_edges(hairball, which(E(hairball)$year == 1958)), 
                  mode = 'in')
V(hairball)$pop_devel <- ifelse(pop1957 < pop1958, 'increased',
                                ifelse(pop1957 > pop1958, 'decreased', 
                                       'unchanged'))
V(hairball)$popularity <- pmax(pop1957, pop1958)
E(hairball)$year <- as.character(E(hairball)$year)
```
    
##Elbow
###Heirachy is the data frame name we want to reference   
```{r, fig.height =3, fig.width =3}
ggraph(hierarchy, layout = 'dendrogram') + 
    geom_edge_elbow()
```

##Heirachy
```{r, fig.height =3, fig.width =3}
graph <- graph_from_data_frame(flare$edges, vertices = flare$vertices)
set.seed(1)
ggraph(graph, 'circlepack', weight = 'size') + 
    geom_node_circle(aes(fill = depth), size = 0.25, n = 50) + 
    coord_fixed()
```


##2-Variant
```{r, fig.height =4, fig.width =4}
ggraph(hierarchy, layout = 'dendrogram') + 
    geom_edge_elbow2(aes(colour = node.class))
```

##Diagnols
```{r, fig.height =3, fig.width =3}
ggraph(hierarchy, layout = 'dendrogram') + 
    geom_edge_diagonal()
```
##########################################   
   
#Geomnet
The net geom is used visualize networks within the ggplot2 framework. geom_net combines the many parts of a network visualization into a single layer in ggplot2. It makes use of various other geoms, including but not limited to, `geom_point`, `geom_segment`, and `geom_text`

```{r}
# Not run: 
library(geomnet)
data(blood)
p <- ggplot(data = blood$edges, aes(from_id = from, to_id = to))
p + geom_net(vertices=blood$vertices, aes(colour=..type..)) + theme_net()

bloodnet <- fortify(as.edgedf(blood$edges), blood$vertices)
p <- ggplot(data = bloodnet, aes(from_id = from_id, to_id = to_id))
p + geom_net()
p + geom_net(aes(colour=rho)) + theme_net()
p + geom_net(aes(colour=rho), labelon=TRUE, vjust = -0.5)
p + geom_net(aes(colour=rho, linetype = group_to, label = from_id),  
             vjust=-0.5, labelcolour="black", directed=TRUE) + 
     theme_net()
p + geom_net(colour = "orange", layout.alg = 'circle', size = 6)
p + geom_net(colour = "orange", layout.alg = 'circle', size = 6, linewidth=.75)
p + geom_net(colour = "orange", layout.alg = 'circle', size = 0, linewidth=.75,
             directed = TRUE)
p + geom_net(aes(size=Predominance, colour=rho, shape=rho, linetype=group_to),
             linewidth=0.75, labelon =TRUE, labelcolour="black") +
    facet_wrap(~Ethnicity) +
    scale_colour_brewer(palette="Set2")
gg <- ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(colour = "darkred", layout.alg = "circle", labelon=TRUE, size = 15,
         directed = TRUE, vjust = 0.5, labelcolour = "grey80",
         arrowsize = 1.5, linewidth = 0.5, arrowgap = 0.05,
         selfloops = TRUE, ecolour = "grey40") +
  theme_net()
gg
dframe <- ggplot_build(gg)$data[[1]] # contains calculated node and edge values

```
```{r}
## visualizing ggplot2 theme elements
data(theme_elements)
TEnet <- fortify(as.edgedf(theme_elements$edges[,c(2,1)]), theme_elements$vertices)
ggplot(data = TEnet, aes(from_id = from_id, to_id = to_id)) +
  geom_net(labelon=TRUE, vjust=-0.5)
```


```{r}
## emails example from VastChallenge 2014
# care has to be taken to make sure that for each panel all nodes are included with
# the necessary information.
# Otherwise line segments show on the plot without nodes.
emailedges <- as.edgedf(subset(email$edges, nrecipients < 54))
emailnet <- fortify(emailedges, email$nodes)
#no facets
ggplot(data = emailnet, aes(from_id = from_id, to_id = to_id)) +
  geom_net(aes(colour= CurrentEmploymentType), linewidth=0.5) +
  scale_colour_brewer(palette="Set2")
#facet by day

emailnet <- fortify(emailedges, email$nodes, group = "day")
ggplot(data = emailnet, aes(from_id = from, to_id = to_id)) +
  geom_net(aes(colour= CurrentEmploymentType), linewidth=0.5, fiteach=TRUE) +
  scale_colour_brewer(palette="Set2") +
  facet_wrap(~day, nrow=2) + theme(legend.position="bottom")
ggplot(data = emailnet, aes(from_id = from, to_id = to_id)) +
  geom_net(aes(colour= CitizenshipCountry), linewidth=0.5, fiteach=TRUE) +
  scale_colour_brewer(palette="Set2") +
  facet_wrap(~day, nrow=2) + theme(legend.position="bottom")
ggplot(data = emailnet, aes(from_id = from, to_id = to_id)) +
  geom_net(aes(colour= CurrentEmploymentType), linewidth=0.5, fiteach=FALSE) +
  scale_colour_brewer(palette="Set2") +
  facet_wrap(~day, nrow=2) + theme(legend.position="bottom")

```
    
    
    
    ```{r}
#Madmen Relationships
data(madmen)
MMnet <- fortify(as.edgedf(madmen$edges), madmen$vertices)
p <- ggplot(data = MMnet, aes(from_id = from_id, to_id = to_id))
p + geom_net(labelon=TRUE)
p + geom_net(aes(colour=Gender), size=6, linewidth=1, labelon=TRUE, fontsize=3, labelcolour="black")
p + geom_net(aes(colour=Gender), size=6, linewidth=1, labelon=TRUE, labelcolour="black") +
    scale_colour_manual(values=c("#FF69B4", "#0099ff")) + xlim(c(-.05,1.05))
p + geom_net(aes(colour=Gender), size=6, linewidth=1, directed=TRUE, labelon=TRUE,
             arrowgap=0.01, labelcolour="black") +
    scale_colour_manual(values=c("#FF69B4", "#0099ff")) + xlim(c(-.05,1.05))

p <- ggplot(data = MMnet, aes(from_id = from_id, to_id = to_id))
# alternative labelling: specify label aesthetic.
p + geom_net(aes(colour=Gender, label=Gender), size=6, linewidth=1, fontsize=3,
             labelcolour="black")
```
   
   
   
   
```{r}

## Les Miserables example
data(lesmis)
lesmisnet <- fortify(as.edgedf(lesmis$edges), lesmis$vertices[, c(2,1)])
p <- ggplot(data=lesmisnet, aes(from_id=from_id, to_id=to_id))
p + geom_net(layout.alg="fruchtermanreingold")
p + geom_net(layout.alg="fruchtermanreingold", labelon=TRUE, vjust=-0.5)
p + geom_net(layout.alg="fruchtermanreingold", labelon=TRUE, vjust=-0.5,
    aes(linewidth=degree/5))

```
   
   
```{r}

## College Football Games in the Fall 2000 regular season
# Source: http://www-/personal.umich.edu/~mejn/netdata/
data(football)
ftnet <- fortify(as.edgedf(football$edges), football$vertices)
p <- ggplot(data=ftnet, aes(from_id=from_id, to_id=to_id))
p + geom_net(aes(colour=value), linewidth=0.75, size=4.5, ecolour="grey80") +
  scale_colour_brewer("Conference", palette="Paired") + theme_net() +
  theme(legend.position="bottom")
  
```
