---
title: "Colleen_heirarchical"
author: "Colleen Brents"
date: "November 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#colleen's to do
+ help files for 'hclust'
rplot 
scale columns (subtract mean and divide by SD) before heirarchical clustering
+ look at scale function to get idea 
+ might want to use mutate_all to scale all columns before run `as.matrix` 
+ ggraph and geomnet 
+ principle components analysis (second)
+ alluvial (which variables could be treated categorical) (third)

reference example_ggraph 
#split by chemical and biological markers 
*Colleen does chemical markers 

```{r}
library(readr)
library(dplyr)
library(ggraph)

test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
                             "master/CSV_Files/efficacy_summary.csv"))

test_data_dend <- test_data %>% 
  select(cLogP:MacUptake) %>% 
  as.matrix() %>% 
  t() %>% 
  dist() %>% 
  hclust() %>% 
  as.dendrogram()

ggraph(test_data_dend, layout = "dendrogram") + 
  geom_edge_diagonal()
```

