---
title: "Colleen_heirarchical"
author: "Colleen Brents"
date: "November 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#colleen's to do
+ help files for 'hclust'
[hclust](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/hclust.html)
rplot (alluvial diagrams stuff)
scale columns (subtract mean and divide by SD) before heirarchical clustering
+ look at scale function to get idea 
+ might want to use mutate_all to scale all columns before run `as.matrix` 
+ ggraph and geomnet 
[ggraph](https://cran.r-project.org/web/packages/ggraph/index.html) 
check out vingettes for edges,  layouts, nodes 
+ principle components analysis (second)
+ alluvial (which variables could be treated categorical) (third)

reference example_ggraph 
#split by chemical and biological markers 
*Colleen does chemical markers 

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggraph)
library(tibble)

test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
                             "master/CSV_Files/efficacy_summary.csv"))

test_data_dend <- test_data %>% 
  unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
  remove_rownames %>% 
  column_to_rownames (var = "drugz") %>%  #make drugz row name 
  select(PLA:MacUptake, -ELU, -ESP) %>%
  mutate_each_(funs(scale(.) %>% as.vector),
               vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "cLogP", "huPPB","muPPB",
               "MIC_Erdman", "MICserumErd", "Caseum_binding", "MacUptake")) 
#mutate_at(funs(scale(.) %>% as.vector), vars)   #mutate_at can be used to scale certain columns (keep the unite identifities)

test_data_matrix<- test_data_dend %>% 
  as.matrix() %>%  #make matrix 
   #filter by drug dosage and then by drug interval 
   t() %>%  #transpose function 
  dist() %>% 
  hclust() %>% 
  #want to add labels (to nodes)
  as.dendrogram()
  
  #set drug:level as matrix row names
  #look at clustering across drugs or across measurements 
  #colleen is doing across drugs


ggraph(test_data_matrix, layout = "dendrogram") + 
  geom_edge_diagonal()
```

