---
title: "Colleen_heirarchical"
author: "Colleen Brents"
date: "November 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#colleen's to do
+ help files for 'hclust'
[hclust](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/hclust.html)
rplot (alluvial diagrams stuff)
scale columns (subtract mean and divide by SD) before heirarchical clustering
+ look at scale function to get idea 
+ might want to use mutate_all to scale all columns before run `as.matrix` 
+ ggraph and geomnet 
[ggraph](https://cran.r-project.org/web/packages/ggraph/index.html) 
check out vingettes for edges,  layouts, nodes 
+ principle components analysis (second)
+ alluvial (which variables could be treated categorical) (third)

reference example_ggraph 
#split by chemical and biological markers 
*Colleen does chemical markers 
[tutorial](http://www.sthda.com/english/articles/28-hierarchical-clustering-essentials/92-visualizing-dendrograms-ultimate-guide/)

```{r}
library(readr)
library(dplyr)
library(tidyr) #unite
library(ggraph)
library(tibble)
library(dendextend)
```

```{r}
# test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
#                              "master/CSV_Files/efficacy_summary.csv"))
# 
# test_data_dend <- test_data %>% 
#   tidyr::unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
#   remove_rownames %>% 
#   column_to_rownames (var = "drugz") %>%  #make drugz row name 
#   select(PLA:MacUptake, -ELU, -ESP) %>%
#   mutate_each_(funs(scale(.) %>% as.vector),
#                vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE", "cLogP", "huPPB","muPPB",
#                "MIC_Erdman", "MICserumErd", "Caseum_binding", "MacUptake")) %>% 
#   as.matrix() %>% 
#   t() %>% 
#   dist() %>% 
#   hclust() %>% 
#   as.dendrogram(horiz = TRUE, hang = .1)
# 
# plot (test_data_dend,
#      xlab = "",
#      ylab = "",
#      main = "Across Drugz Measurements",
#      horiz = TRUE,
#      axes = FALSE)
#      
# par(cex = 0.6, mar=c(9,11,7,7)) #cex magnifies text; mar does axis 
# par(cex = 0.6)
# 
# #color for chemical vs biological 
```

```{r add colors}
test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
                             "master/CSV_Files/efficacy_summary.csv"))

test_data_dend <- test_data %>% 
  tidyr::unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
  remove_rownames %>% 
  column_to_rownames (var = "drugz") %>%  #make drugz row name 
  select(PLA:MacUptake, -ELU, -ESP) %>%
  mutate_each_(funs(scale(.) %>% as.vector),
               vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE", "cLogP", "huPPB","muPPB",
               "MIC_Erdman", "MICserumErd", "MIC_Rv", "Caseum_binding", "MacUptake")) %>% 
  as.matrix() %>% 
  t() %>% 
  dist() %>% 
  hclust() %>% 
  as.dendrogram(horiz = TRUE, hang = .1) %>% 
  #add mackenzie's
  #plot.dendorgram help file 
  plot(xlab = "",
     ylab = "",
     main = "Across Drug Measurements",
     horiz = TRUE,
     axes = FALSE) %>% 
  set("labels")
  set("labels", c(111:115)) %>%    # change labels
   set("labels_col", c(1,2,3)) %>%  # change color 
  par(cex = 0.6, mar=c(9,11,7,7)) %>% #cex magnifies text; mar does axis
  par(cex = 0.6)
#add column 
invivo_bio<- c("cLogP", "huPPB","muPPB", "MIC_Erdman", "MICserumErd", "MIC_Rv","Caseum_binding", "MacUptake") 
invitro_chem<- c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE")

#color for chemical vs biological indepdent of the cluster; different grouping 
```

```{r}
dend15 <- c(1:5) %>% dist %>% hclust(method = "average") %>% as.dendrogram
dend15 %>% set("labels_to_char") %>% labels

dend15_2 <- dend15 %>% 
   set("labels", c(111:115)) %>%    # change labels
   set("labels_col", c(1,2,3)) %>%  # change color 
   set("labels_cex", c(2,1))
par(mfrow = c(1,2))
dend15 %>% plot(main = "Before")
dend15_2 %>% plot(main = "After")
```

the above graph visualizes how similar test results are across all drug-level-trt combinations 

resource for color coding leaves [here](https://stackoverflow.com/questions/31117849/how-to-color-a-dendrograms-labels-according-to-defined-groups-in-r)
learn about extendogram [here](https://cran.r-project.org/web/packages/dendextend/vignettes/introduction.html)
also [here](https://rpubs.com/gaston/dendrograms) 
```{r data from R for colors}
fviz_dend(test_data_dend, k = 6, # Cut in four groups
          cex = 0.5,
          xlab = "",
          ylab = "",
          main = "Cluster by Measurements",
          horiz = TRUE,
          k_colors = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07", "#F09428",
                       "#18E154"),
          color_labels_by_k = FALSE, # color labels by groups
          rect = TRUE, # Add rectangle around groups
          rect_border = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07", "#F09428",
                          "#18E154"),
          rect_fill = TRUE)
```


```{r}
test_data_dend_drug <- test_data %>% 
  tidyr::unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
  remove_rownames %>% 
  column_to_rownames (var = "drugz") %>%  #make drugz row name 
  select(PLA:MacUptake, -ELU, -ESP) %>%
  #keep column names
  #pull out drug names and save as vector 
  
  mutate_each_(funs(scale(.) %>% as.vector),
               vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE", "cLogP", "huPPB","muPPB",
               "MIC_Erdman", "MICserumErd", "Caseum_binding", "MacUptake")) %>% 
  mutate_each_(funs(scale(.)),
               vars ("PLA":"MacUptake"))  #<- works like select 
  as.matrix() %>% 
  dist() %>% 
  hclust() %>% 
  as.dendrogram(horiz = TRUE, hang = .1)

plot (test_data_dend_drug,
     xlab = "",
     ylab = "",
     main = "Across Drugz Measurements",
     horiz = TRUE,
     axes = FALSE)
     
par(cex = 0.6, mar=c(9,11,7,7)) #cex magnifies text; mar does axis 
par(cex = 0.6)

#make labels names as a vector, add back later 
#go step by step ; col names of matrix = 
```

