---
title: "Colleen_heirarchical"
author: "Colleen Brents"
date: "November 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#colleen's to do
+ help files for 'hclust'
[hclust](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/hclust.html)
rplot (alluvial diagrams stuff)
scale columns (subtract mean and divide by SD) before heirarchical clustering
+ look at scale function to get idea 
+ might want to use mutate_all to scale all columns before run `as.matrix` 
+ ggraph and geomnet 
[ggraph](https://cran.r-project.org/web/packages/ggraph/index.html) 
check out vingettes for edges,  layouts, nodes 
+ principle components analysis (second)
+ alluvial (which variables could be treated categorical) (third)

reference example_ggraph 
#split by chemical and biological markers 
*Colleen does chemical markers 
[tutorial](http://www.sthda.com/english/articles/28-hierarchical-clustering-essentials/92-visualizing-dendrograms-ultimate-guide/)

```{r}
library(readr)
library(dplyr)
library(tidyr) #unite
library(ggraph)
library(tibble)
library(dendextend)
```

```{r add colors}
test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
                             "master/CSV_Files/efficacy_summary.csv"))

test_data_dend <- test_data %>% 
  tidyr::unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
  remove_rownames %>% 
  column_to_rownames (var = "drugz") %>%  #make drugz row name 
  select(PLA:MacUptake, -ELU, -ESP) %>%  #remove efficacy 
  mutate_each_(funs(scale(.) %>% as.vector),
               vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE", "cLogP", "huPPB","muPPB",
               "MIC_Erdman", "MICserumErd", "MIC_Rv", "Caseum_binding", "MacUptake")) %>% 
  as.matrix() %>% 
  t() %>% #transpose 
  dist() %>% 
  hclust() %>% 
  as.dendrogram(horiz = TRUE, hang = .1) %>% 
  plot(xlab = "",
     ylab = "",
     main = "Across Drug Measurements",
     horiz = TRUE,
     axes = FALSE)
  par(cex = 0.6, mar=c(9,11,7,7)) %>% #cex magnifies text; mar does axis
  par(cex = 0.6)
#add column 
invivo_bio<- c("cLogP", "huPPB","muPPB", "MIC_Erdman", "MICserumErd", "MIC_Rv","Caseum_binding", "MacUptake") 
invitro_chem<- c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE")

#color for chemical vs biological indepdent of the cluster; different grouping 
# differenet heirarchical algorithms 
# add invitro and invivo identifiers 
# add full names 

```

RIM = rim
OCS - outer caeseum
ICS -inner caseesum
uninvolved lung ULU
ULU -> rim -> OCS --> ICS  
pla plasma
sle - standard lesion (entire)
standard lung SLU 
mic - minimum inhibiroy concenration 


resource for color coding leaves [here](https://stackoverflow.com/questions/31117849/how-to-color-a-dendrograms-labels-according-to-defined-groups-in-r)
learn about extendogram [here](https://cran.r-project.org/web/packages/dendextend/vignettes/introduction.html)
also [here](https://rpubs.com/gaston/dendrograms) 

```{r now categorize invivo invitro}
test_data <- read_csv(paste0("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/",
                             "master/CSV_Files/efficacy_summary.csv"))

sortr <- test_data %>% 
  tidyr::unite(drugz, drug:level, sep = "_") %>% #combine identifying data into one column, 
  remove_rownames %>% 
  column_to_rownames (var = "drugz") %>%  #make drugz row name 
  select(PLA:MacUptake, -ELU, -ESP) %>%  #remove efficacy 
  mutate_each_(funs(scale(.) %>% as.vector),
               vars = c("PLA", "ULU", "RIM", "OCS", "ICS", "SLU", "SLE", "cLogP", "huPPB","muPPB",
               "MIC_Erdman", "MICserumErd", "MIC_Rv", "Caseum_binding", "MacUptake")) %>% 
  t() %>% 
  #add column to identify invivo or invitro
  #add identifier to invivo or invitro 
  mutate(invivo = c("cLogP", "huPPB","muPPB", "MIC_Erdman", "MICserumErd", "MIC_Rv","Caseum_binding", "MacUptake"))
  as.matrix() %>% 
  t() #transpose 
  as.data.
  
  sortr<- sortr %>% 
    as.data.frame() %>% 
    column_to_rownames(var = "invivo")
    mutate(invivo, )

```





